Policy Gradient Estimation

### 2019-03-03 pge-mattis.tex
Sentences in separate lines

Policy gradient methods are approaches to maximize the expected return in a Markov Decision Process (MDP). Using a parameterized policy to decide the next action, they can easily incorporate prior domain knowledge, but require a lot of configuration to produce an effective agent for a specific environment. Also, they frequently require on-policy training and often converge slower, but can guaranteed find a local optimum given some constraints on the hyperparameters. Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed. \cite{DBLP:journals/corr/SchulmanMLJA15} In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e. improve its expected return. Policy gradient estimation is the term we use to describe the process of computing the next step in parameter space. The goal is to estimate the gradient of the policy with respect to the expected return. Since this is the core problem of policy gradient methods, it is also the main topic of this paper.

@@

\paragraph{Policy gradient.} Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$, also called expected return. We define $p(s|\theta)$ as the probability The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$. We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended. Intuitively, this reflects the idea that the relevance of later actions declines, which puts more weight on less actions, thus the increased bias for reduced variance. The discounted total reward is

@@

Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities. Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. This means that actions can only become less probable in relation to other actions. We find that this introduces a lot of variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. One approach to counter the variance is to introduce a baseline
\begin{equation}
  \psi_t \leftarrow \gamma^{t} (r_{t} - b(s_t))
\end{equation}
The baseline can represent any prior knowledge or assumption about states, e.g., their mean value. Though we find that there is a specific way to use this baseline which fits surprisingly well into this framework.

>>>>

Policy gradient methods are approaches to maximize the expected return in a Markov Decision Process (MDP). 
Using a parameterized policy to decide the next action, they can easily incorporate prior domain knowledge, but require a lot of configuration to produce an effective agent for a specific environment. 
Also, they frequently require on-policy training and often converge slower, but can guaranteed find a local optimum given some constraints on the hyperparameters. 
Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed \cite{DBLP:journals/corr/SchulmanMLJA15}. 
In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e. improve its expected return. 
Policy gradient estimation is the term we use to describe the process of computing the next step in parameter space. 
The goal is to estimate the gradient of the policy with respect to the expected return. 
Since this is the core problem of policy gradient methods, it is also the main topic of this paper.

@@

\paragraph{Policy gradient.} 
Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$, also called expected return. 
We define $p(s|\theta)$ as the probability TODO
The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$. 
We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended.
Intuitively, this reflects the idea that the relevance of later actions declines, which puts more weight on less actions, thus the increased bias for reduced variance. 
The discounted total reward is

@@

Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities. 
Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
This means actions can only become less probable in relation to other actions. 
We find that this introduces a lot of variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
One approach to counter the variance is to introduce a baseline
\begin{equation}
  \psi_t \leftarrow \gamma^{t} (r_{t} - b(s_t))
\end{equation}
The baseline can represent any prior knowledge or assumption about states, e.g., their mean value. 
Though we find that there is a specific way to use this baseline which fits surprisingly well into this framework.
